---
title: "Rice-Wheat System Spatial Decision Support Tool"
format: 
    html: 
      code-fold: true
      code-tools: true
fig-dpi: 300
fig-width: 8.88
fig-align: center
fig-height: 5

self-contained: true
author: Maxwell Mkondiwa (m.mkondiwa@cgiar.org)

editor: visual
toc: true
toc-location: left
number-sections: true
execute: 
  message: false
  warning: false
  echo: true
---

# Introduction

We present a rice-wheat systen spatial decisionb support tool which relies on data from the matched landscape crop assessment surveys in eastern India and a multivariate geoadditive Bayesian model to predict entry points for system optimization in the area of interest.

# Crop specific regressions

## Rice

```{r}
library(sp)
library(mgcv)
library(bamlss)

# Multivariate geoadditive model
# remotes::install_git("https://git.uibk.ac.at/c4031039/mvnchol")
# library(mvnchol)
library(BayesX)
library(R2BayesX)
library(sf)
library(spdep)
library(rio)
Irrig_Rev_rice_wheat <- import("data/Irrig_Rev_rice_wheat.csv")

library(janitor)
# Irrig_Rev_rice_wheat <- clean_names(Irrig_Rev_rice_wheat)
library(tidyr)
#Irrig_Rev_rice_wheat <- Irrig_Rev_rice_wheat %>% drop_na()

# library(lubridate)
# library(anytime)
# Irrig_Rev_rice_wheat$januaryfirst2017 <- ymd("2017-01-01")

# Irrig_Rev_rice_wheat$sowdate_fmt_wheat_day <- Irrig_Rev_rice_wheat$sowdate_fmt - Irrig_Rev_rice_wheat$januaryfirst2017

# Irrig_Rev_rice_wheat$sowdate_fmt_rice_day <- Irrig_Rev_rice_wheat$sowdate_fmt_rice - Irrig_Rev_rice_wheat$januaryfirst2017

Irrig_Rev_rice_wheat$harvest_day_rice <- Irrig_Rev_rice_wheat$l_crop_duration_days_rice + Irrig_Rev_rice_wheat$sowdate_fmt_rice_day

# Irrig_Rev_rice_wheat$harvest_day_wheat <- Irrig_Rev_rice_wheat$l_crop_duration_days + Irrig_Rev_rice_wheat$sowdate_fmt_wheat_day


shpname <- file.path(getwd(), "shp", "India_aoi_sf_sp")
India_aoi_sp_bnd <- BayesX::shp2bnd(shpname = shpname, regionnames = "District", check.is.in = F)

f_rice_yield_MRF <- list(
    b_grain_yield_ton_per_ha_rice ~ 1 + rice_duration_class_long + s(sowdate_fmt_rice_day) + s(g_q5305_irrig_times_rice) + s(nperha_rice) + s(p2o5perha_rice) + s(District, bs = "mrf", xt = list("penalty" = K)) +
        s(District, bs = "re"),
    sigma ~ 1 + rice_duration_class_long + s(sowdate_fmt_rice_day) + s(g_q5305_irrig_times_rice) + s(nperha_rice) + s(p2o5perha_rice) + s(District, bs = "mrf", xt = list("penalty" = K)) +
        s(District, bs = "re")
)



K <- neighbormatrix(India_aoi_sp_bnd)
head(K)

## Also need to transform to factor for
## setting up the MRF smooth.
Irrig_Rev_rice_wheat$District <- as.factor(Irrig_Rev_rice_wheat$a_q103_district)

## Now note that not all regions are observed,
## therefore we need to remove those regions
## from the penalty matrix
rn <- rownames(K)
lv <- levels(Irrig_Rev_rice_wheat$District)
i <- rn %in% lv
K <- K[i, i]

set.seed(321)
library(bamlss)
b_rice_yield_MRF <- bamlss(f_rice_yield_MRF, data = Irrig_Rev_rice_wheat, family = "gaussian")

## First, note that we have the structured id = 'mrf1' and unstructured
## spatial effect id = 're2', also indicated in the model summary
summary(b_rice_yield_MRF)

# Plot the nonlinear effect
plot(b_rice_yield_MRF, model = "mu", term = "s(sowdate_fmt_rice_day)")
plot(b_rice_yield_MRF, model = "mu", term = "s(g_q5305_irrig_times_rice)")
plot(b_rice_yield_MRF, model = "mu", term = "s(nperha_rice)")
plot(b_rice_yield_MRF, model = "mu", term = "s(p2o5perha_rice)")





```

## Wheat model

```{r}
f_wheat_yield_MRF <- list(
    l_ton_per_hectare ~ 1 + variety_type_NMWV + s(sowdate_fmt_wheat_day) + g_q5305_irrig_times + nperha + p2o5perha + s(District, bs = "mrf", xt = list("penalty" = K)) +
        s(District, bs = "re"),
    sigma ~ 1 + variety_type_NMWV + s(sowdate_fmt_wheat_day) + g_q5305_irrig_times + nperha + p2o5perha + s(District, bs = "mrf", xt = list("penalty" = K)) +
        s(District, bs = "re")
)
set.seed(321)
b_wheat_yield_MRF <- bamlss(f_wheat_yield_MRF, data = Irrig_Rev_rice_wheat, family = "gaussian")

## First, note that we have the structured id = 'mrf1' and unstructured
## spatial effect id = 're2', also indicated in the model summary
summary(b_wheat_yield_MRF)

# Plot the nonlinear effect
plot(b_wheat_yield_MRF, model = "mu", term = "s(sowdate_fmt_wheat_day)")

## Now, to predict the spatial effects we set up new data.
nd <- data.frame("District" = unique(Irrig_Rev_rice_wheat$District))

## Predict for the structured spatial effects.
p_str_wheat_yield_MRF <- predict(b_wheat_yield_MRF, newdata = nd, term = "s(District,id='mrf1')", intercept = FALSE)

## And the unstructured spatial effect.
p_unstr_wheat_yield_MRF <- predict(b_wheat_yield_MRF, newdata = nd, term = "s(District,id='re2')", intercept = FALSE)

## MRF smooth effect.
plotmap(India_aoi_sp_bnd,
    x = p_str_wheat_yield_MRF$mu, id = nd$District,
    main = expression(mu), title = "Structured spatial effect"
)
plotmap(India_aoi_sp_bnd,
    x = p_str_wheat_yield_MRF$sigma, id = nd$District,
    main = expression(sigma), title = "Structured spatial effect"
)

## Random effects.
plotmap(India_aoi_sp_bnd,
    x = p_unstr_wheat_yield_MRF$mu, id = nd$District, title = "Unstructured spatial effect"
)
plotmap(India_aoi_sp_bnd,
    x = p_unstr_wheat_yield_MRF$sigma, id = nd$District, title = "Unstructured spatial effect"
)



```

# Multivariate Spatial Geoadditive Bayesian Regression Model

# Spatial + for multivariate model

```{r}
library(sp)
library(mgcv)
library(bamlss)

# Multivariate geoadditive model
# remotes::install_git("https://git.uibk.ac.at/c4031039/mvnchol")
# install_github("https://github.com/meteosimon/mvnchol")
library(mvnchol)

# Remove NAs in the monsoon variables
Irrig_Rev_rice_wheat <- subset(Irrig_Rev_rice_wheat, !(is.na(Irrig_Rev_rice_wheat$onset_2017)))
Irrig_Rev_rice_wheat <- subset(Irrig_Rev_rice_wheat, !(is.na(Irrig_Rev_rice_wheat$monsoon_onset_dev)))
Irrig_Rev_rice_wheat <- subset(Irrig_Rev_rice_wheat, !(is.na(Irrig_Rev_rice_wheat$median_onset_82_15)))
Irrig_Rev_rice_wheat <- subset(Irrig_Rev_rice_wheat, !(is.na(Irrig_Rev_rice_wheat$sd_onset_82_15)))

# Rice first stage
f_sow_rice_1st_stage <- list(
    sowdate_fmt_rice_day ~ 1 + rice_duration_class_long + s(gw_2017) + s(onset_2017) + s(monsoon_onset_dev) + s(median_onset_82_15) + s(sd_onset_82_15) + s(District, bs = "mrf", xt = list("penalty" = K)) +
        s(District, bs = "re")
)

f_sow_rice_1st_stage_MRF <- bamlss(f_sow_rice_1st_stage, data = Irrig_Rev_rice_wheat, family = "gaussian")

fitted_f_sow_rice_1st_stage_MRF <- f_sow_rice_1st_stage_MRF$fitted

Irrig_Rev_rice_wheat$Res_rice_sow <- Irrig_Rev_rice_wheat$sowdate_fmt_rice_day - fitted_f_sow_rice_1st_stage_MRF$mu

# Wheat first stage
f_sow_wheat_1st_stage <- list(
    sowdate_fmt_wheat_day ~ 1 + variety_type_NMWV + s(harvest_day_rice) + s(gw_2018) + s(District, bs = "mrf", xt = list("penalty" = K)) +
        s(District, bs = "re")
)

f_sow_wheat_1st_stage_MRF <- bamlss(f_sow_wheat_1st_stage, data = Irrig_Rev_rice_wheat, family = "gaussian")

fitted_f_sow_wheat_1st_stage_MRF <- f_sow_wheat_1st_stage_MRF$fitted

Irrig_Rev_rice_wheat$Res_wheat_sow <- Irrig_Rev_rice_wheat$sowdate_fmt_wheat_day - fitted_f_sow_wheat_1st_stage_MRF$mu


# Multivariate with sowing dates as endogenous
f_rice_wheat_yield_MRF_corr_endo <- list(
    sowdate_fmt_rice_day ~ 1 + rice_duration_class_long + s(gw_2017) + s(onset_2017) + s(monsoon_onset_dev) + s(median_onset_82_15) + s(sd_onset_82_15) + s(District, bs = "mrf", xt = list("penalty" = K)) +
        s(District, bs = "re"),
    b_grain_yield_ton_per_ha_rice ~ 1 + rice_duration_class_long + s(Res_rice_sow) + s(g_q5305_irrig_times_rice) + s(nperha_rice) + s(p2o5perha_rice) + s(District, bs = "mrf", xt = list("penalty" = K)) +
        s(District, bs = "re"),
    sowdate_fmt_wheat_day ~ 1 + variety_type_NMWV + s(harvest_day_rice) + s(District, bs = "mrf", xt = list("penalty" = K)) +
        s(District, bs = "re"),
    l_ton_per_hectare ~ 1 + variety_type_NMWV + s(Res_wheat_sow) + g_q5305_irrig_times + nperha + p2o5perha + s(District, bs = "mrf", xt = list("penalty" = K)) + s(District, bs = "re"),
    lamdiag1 ~ 1,
    lamdiag2 ~ 1,
    lamdiag3 ~ 1,
    lamdiag4 ~ 1,
    lambda12 ~ 1,
    lambda13 ~ 1,
    lambda14 ~ 1,
    lambda23 ~ 1,
    lambda24 ~ s(District, bs = "mrf", xt = list("penalty" = K)) + s(District, bs = "re"),
    lambda34 ~ 1
)

multivariate_geo_sow_MRF_corr_endo <- bamlss(f_rice_wheat_yield_MRF_corr_endo, type = "modified", family = mvnchol_bamlss(k = 4), data = Irrig_Rev_rice_wheat)

summary(multivariate_geo_sow_MRF_corr_endo)

nd <- data.frame("District" = unique(Irrig_Rev_rice_wheat$District))

# Focusing on the cross-equation correlations

## Predict for the structured spatial effects.
p_str_multivariate_geo_sow_MRF_corr_endo_rice_y <- predict(multivariate_geo_sow_MRF_corr_endo, newdata = nd, term = "s(District,id='mrf1')", intercept = FALSE)

p_str_multivariate_geo_sow_MRF_corr_endo_rice_ydt <- as.data.frame(p_str_multivariate_geo_sow_MRF_corr_endo_rice_y)


## And the unstructured spatial effect.
p_unstr_multivariate_geo_sow_MRF_corr_endo_rice_y <- predict(multivariate_geo_sow_MRF_corr_endo, newdata = nd, term = "s(District,id='re2')", intercept = FALSE)

# Rice sowing spatial equation

plotmap(India_aoi_sp_bnd, x = p_str_multivariate_geo_sow_MRF_corr_endo_rice_y$mu1, id = nd$District, main = expression(mu(rice_sowing)), title = "Structured spatial effect")

plotmap(India_aoi_sp_bnd, x = p_unstr_multivariate_geo_sow_MRF_corr_endo_rice_y$mu1, id = nd$District, main = expression(mu(rice_sowing)), title = "Unstructured spatial effect")

# Rice yield spatial equation
plotmap(India_aoi_sp_bnd, x = p_str_multivariate_geo_sow_MRF_corr_endo_rice_y$mu2, id = nd$District, main = expression(mu(rice_yield)), title = "Structured spatial effect")

plotmap(India_aoi_sp_bnd, x = p_unstr_multivariate_geo_sow_MRF_corr_endo_rice_y$mu2, id = nd$District, main = expression(mu(rice_yield)), title = "Unstructured spatial effect")

# Wheat sowing equation
plotmap(India_aoi_sp_bnd, x = p_str_multivariate_geo_sow_MRF_corr_endo_rice_y$mu3, id = nd$District, main = expression(mu(wheat_sowing)), title = "Structured spatial effect")

plotmap(India_aoi_sp_bnd, x = p_unstr_multivariate_geo_sow_MRF_corr_endo_rice_y$mu3, id = nd$District, main = expression(mu(wheat_sowing)), title = "Unstructured spatial effect")

# Wheat yield spatial equation
plotmap(India_aoi_sp_bnd, x = p_str_multivariate_geo_sow_MRF_corr_endo_rice_y$mu4, id = nd$District, main = expression(mu(wheat_yield)), title = "Structured spatial effect")

plotmap(India_aoi_sp_bnd, x = p_unstr_multivariate_geo_sow_MRF_corr_endo_rice_y$mu4, id = nd$District, main = expression(mu(wheat_yield)), title = "Unstructured spatial effect")

# Focusing on the cross-equation correlations

## MRF smooth effect.
plotmap(India_aoi_sp_bnd,
    x = p_str_multivariate_geo_sow_MRF_corr_endo_rice_y$lambda24, id = nd$District,
    main = expression(lambda(rice, wheat)), title = "Structured spatial effect"
)

## Random effects.
plotmap(India_aoi_sp_bnd,
    x = p_unstr_multivariate_geo_sow_MRF_corr_endo_rice_y$lambda24, id = nd$District, main = expression(lambda(rice, wheat)), title = "Unstructured spatial effect"
)






# Rice sowing equation : Non-linear relationships
# s(gw_2017) + s(onset_2017) + s(monsoon_onset_dev) + s(median_onset_82_15) + s(sd_onset_82_15)
plot(multivariate_geo_sow_MRF_corr_endo, model = "mu1", term = "s(gw_2017)")
plot(multivariate_geo_sow_MRF_corr_endo, model = "mu1", term = "s(onset_2017)")
plot(multivariate_geo_sow_MRF_corr_endo, model = "mu1", term = "s(monsoon_onset_dev)")
plot(multivariate_geo_sow_MRF_corr_endo, model = "mu1", term = "s(median_onset_82_15)")
plot(multivariate_geo_sow_MRF_corr_endo, model = "mu1", term = "s(sd_onset_82_15)")


# Rice yield equation
# s(Res_rice_sow) + s(g_q5305_irrig_times_rice) + s(nperha_rice) + s(p2o5perha_rice)

plot(multivariate_geo_sow_MRF_corr_endo, model = "mu2", term = "s(Res_rice_sow)")

plot(multivariate_geo_sow_MRF_corr_endo, model = "mu2", term = "s(g_q5305_irrig_times_rice)")

plot(multivariate_geo_sow_MRF_corr_endo, model = "mu2", term = "s(nperha_rice)")

plot(multivariate_geo_sow_MRF_corr_endo, model = "mu2", term = "s(p2o5perha_rice)")

# Wheat sowing equation
# s(harvest_day_rice) + s(gw_2018)
plot(multivariate_geo_sow_MRF_corr_endo, model = "mu3", term = "s(harvest_day_rice)")

# (multivariate_geo_sow_MRF_corr_endo, model = "mu3", term = "s(gw_2018)")

# Wheat yield equation
plot(multivariate_geo_sow_MRF_corr_endo, model = "mu4", term = "s(Res_wheat_sow)")


# Fitted values
multivariate_geo_sow_MRF_corr_endo_fitted_values <- multivariate_geo_sow_MRF_corr_endo$fitted

multivariate_geo_sow_MRF_corr_endo_fitted_values <- as.data.frame(multivariate_geo_sow_MRF_corr_endo_fitted_values)

# Merge the fitted results to the data and export



Irrig_Rev_rice_wheat_Mult_Res <- cbind(Irrig_Rev_rice_wheat, multivariate_geo_sow_MRF_corr_endo_fitted_values)


summary(lm(sowdate_fmt_rice_day ~ mu1, Irrig_Rev_rice_wheat_Mult_Res))

summary(lm(b_grain_yield_ton_per_ha_rice ~ l_ton_per_hectare, Irrig_Rev_rice_wheat_Mult_Res))



plot((lm(sowdate_fmt_rice_day ~ mu1, Irrig_Rev_rice_wheat_Mult_Res)))

Irrig_Rev_rice_wheat_Mult_Res_sp <- Irrig_Rev_rice_wheat_Mult_Res
coordinates(Irrig_Rev_rice_wheat_Mult_Res_sp) <- c("o_largest_plot_gps_longitude", "o_largest_plot_gps_latitude")

# Map the correlations
# library(modelsummary)
# Irrig_Rev_rice_wheat_Mult_Res_dist <- datasummary(Heading("District") * District ~ Heading("N obs") * N + Heading("%") * Percent() + lambda24 * (Mean + SD), data = Irrig_Rev_rice_wheat_Mult_Res, output = "data.frame")

# library(dplyr)
# Irrig_Rev_rice_wheat_Mult_Res_dist <- rename(Irrig_Rev_rice_wheat_Mult_Res_dist, "Mean_Rice_Wheat_Rho" = "Mean")
# Irrig_Rev_rice_wheat_Mult_Res_dist <- rename(Irrig_Rev_rice_wheat_Mult_Res_dist, "SD_Rice_Wheat_Rho" = "SD")

# Irrig_Rev_rice_wheat_Mult_Res_dist <- subset(Irrig_Rev_rice_wheat_Mult_Res_dist, Irrig_Rev_rice_wheat_Mult_Res_dist$District != "Purnia")

# rice_wheat_yield_rho_dist_sf <- merge(India_aoi_sf, Irrig_Rev_rice_wheat_Mult_Res_dist, by = "District")

# rice_wheat_yield_rho_dist_sf$Mean_Rice_Wheat_Rho <- as.numeric(rice_wheat_yield_rho_dist_sf$Mean_Rice_Wheat_Rho)
# library(mapview)
# mapview(rice_wheat_yield_rho_dist_sf, zcol = "Mean_Rice_Wheat_Rho", layer.name = "Rice wheat equation correlation")
# library(sf)
# rice_wheat_yield_rho_dist_sf_sp <- as_Spatial(rice_wheat_yield_rho_dist_sf)
# library(tmap)
# tmap_mode("view")
# rice_wheat_yield_rho_dist_sf_sp_map <- tm_shape(rice_wheat_yield_rho_dist_sf_sp) +
#     tm_polygons(col = "Mean_Rice_Wheat_Rho", title = "Rice wheat equation correlation", style = "quantile") +
#     tm_layout(legend.outside = TRUE)

# tmap_save(rice_wheat_yield_rho_dist_sf_sp_map, "figures/rice_wheat_yield_rho_dist_sf_sp_map .png")


```

# Factors affecting the correlation structure

```{r}
f_rice_wheat_yield_MRF_corr_endo_General <- list(
    sowdate_fmt_rice_day ~ 1 + rice_duration_class_long + s(gw_2017) + s(onset_2017) + s(monsoon_onset_dev) + s(median_onset_82_15) + s(sd_onset_82_15) + s(District, bs = "mrf", xt = list("penalty" = K)) +
        s(District, bs = "re"),
    b_grain_yield_ton_per_ha_rice ~ 1 + rice_duration_class_long + s(Res_rice_sow) + s(g_q5305_irrig_times_rice) + s(nperha_rice) + s(p2o5perha_rice) + s(District, bs = "mrf", xt = list("penalty" = K)) +
        s(District, bs = "re"),
    sowdate_fmt_wheat_day ~ 1 + variety_type_NMWV + s(harvest_day_rice) + s(District, bs = "mrf", xt = list("penalty" = K)) +
        s(District, bs = "re"),
    l_ton_per_hectare ~ 1 + variety_type_NMWV + s(Res_wheat_sow) + g_q5305_irrig_times + nperha + p2o5perha + s(District, bs = "mrf", xt = list("penalty" = K)) + s(District, bs = "re"),
    lamdiag1 ~ 1,
    lamdiag2 ~ 1,
    lamdiag3 ~ 1,
    lamdiag4 ~ 1,
    lambda12 ~ 1,
    lambda13 ~ 1,
    lambda14 ~ 1,
    lambda23 ~ 1,
    lambda24 ~ s(Res_rice_sow)+s(Res_wheat_sow)+rice_duration_class_long + s(gw_2017) + s(onset_2017)+variety_type_NMWV+s(District, bs = "mrf", xt = list("penalty" = K)) + s(District, bs = "re"),
    lambda34 ~ 1
)

multivariate_geo_sow_MRF_corr_endo_General <- bamlss(f_rice_wheat_yield_MRF_corr_endo_General, type = "modified", family = mvnchol_bamlss(k = 4), data = Irrig_Rev_rice_wheat)

summary(multivariate_geo_sow_MRF_corr_endo_General)

plot(multivariate_geo_sow_MRF_corr_endo_General, model = "lambda24", term = "s(gw_2017)")

plot(multivariate_geo_sow_MRF_corr_endo_General, model = "lambda24", term = "s(onset_2017)")

plot(multivariate_geo_sow_MRF_corr_endo_General, model = "lambda24", term = "s(Res_rice_sow)")

plot(multivariate_geo_sow_MRF_corr_endo_General, model = "lambda24", term = "s(Res_wheat_sow)")

save.image("RW_System_SDS.RData")
```

# Visualization

```{r}
#| context: server
library(distreg.vis)


if (interactive()) {
   distreg.vis:: vis()
}

```

# Conclusion

# References
